---
title: "Big Thompson Data Explore"
author: "Sam Struthers-CSU"
date: "`r Sys.Date()`"
output: html_document
---
# Set Up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(arrow)
library(here)
library(cdssr)
library(tmap)
library(sf)
library(leaflet)
library(tidyverse)
library(nhdplusTools)
library(yaml)
library(snotelr)

knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.dim = c(12,10))
```

Basic overview of preliminary data available in the Big Thompson Basin


# Get Approximate Watershed

```{r}
#Picking a site below the WTP
location_coords <- c(-105.166145, 40.410814)
start_point <- st_sfc(st_point(location_coords), crs = 4326)

start_comid <- discover_nhdplus_id(start_point)

nldi_feature <- list(featureSource = "comid", featureID = start_comid)

#Getting Flowlines and watershed
upstream_nldi <- navigate_nldi(
  nldi_feature = nldi_feature,
  mode = "upstreamTributaries", # "UT" equivalent
  distance_km = 9999
)

big_t_flowlines <- upstream_nldi$UT_flowlines

big_t_watershed <- get_nldi_basin(nldi_feature = nldi_feature)
```


# Read in locations of WQ/Flow Gages (Deployed or Not Yet)

## CDWR Flow Gages

```{r}
creds <- read_yaml(here("creds/CDWRCreds.yml"))$api_key

#Big T is the 4th water district
telemetry <- get_telemetry_stations(water_district = 4, api_key = creds)%>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326)

#clip telemetry to watershed 
telemetry <- telemetry[big_t_watershed,]
```

## Approximate CSU WQ Gages

```{r}
csu_sites <- read.csv("data/csu_sonde_location_metadata.csv")%>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326)

```

## WET WQ Sensors
```{r}

wet_sites <- read.csv("data/wet_sites_big_t.csv")%>%
  #split lat_long into separate columns
  separate(lat_long, into = c("latitude", "longitude"), sep = ",")%>%
  na.omit()%>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326)

```


## Place holder for NW Data

B, is this easy to get and do we want it?

## SNOTEL
```{r}
snotel_sites <- snotel_info()%>%
  filter(grepl("10190006", description, ignore.case = TRUE) | site_name == "lake irene " )%>%
  mutate(`Site Name` = str_to_title(site_name))%>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326)%>%
  select(`Site Name`, site_id, `Start Date` = start, `End Date` = end, geometry)
```


# Map of Sites
```{r}

#Make a map with sites colored by station_type and add the watershed boundary 
tmap_mode("view")

tm_basemap("OpenStreetMap") +
  tm_shape(big_t_watershed) +
  tm_polygons(fill = "lightblue", fill_alpha = 0.5) +
  # flowlines
  tm_shape(big_t_flowlines) +
  tm_lines(col = "blue", lwd = 1) +
  # telemetry stations
  tm_shape(telemetry %>%select(station_type, station_name, abbrev, units, parameter, station_por_start, station_por_end)) +
  tm_dots(
    fill = "station_type",
    size = 1,
    shape = 23,   # filled diamond
    fill.scale = tm_scale_categorical(values = "brewer.set1"),
    fill.legend = tm_legend(title = "CDWR Stations", na.show = FALSE)
  ) +
  # CSU sites
  tm_shape(csu_sites) +
  tm_dots(
    fill = "site_code",
    size = 1,
    shape = 22,   # filled square
    fill.scale = tm_scale_categorical(values = "brewer.set2"),
    fill.legend = tm_legend(title = "CSU Site", na.show = FALSE)
  ) +
  # WET sites
  tm_shape(wet_sites) +
  tm_dots(
    fill = "site_code",
    size = 1,
    shape = 21,   # filled circle
    fill.scale = tm_scale_categorical(values = "brewer.set3"),
    fill.legend = tm_legend(title = "WET Site", na.show = FALSE)
  ) +
  # SNOTEL sites
  tm_shape(snotel_sites) +
  tm_dots(
    fill = "Site Name",
    size = 1,
    shape = 24,   # filled triangle (point-up)
    fill.scale = tm_scale_categorical(values = "Accent"),
    fill.legend = tm_legend(title = "SNOTEL Site", na.show = FALSE)
  ) +
  tm_layout(
    legend.outside = TRUE,
    legend.outside.position = "right"
  )

```

# Code to pull in data from CDWR telemetry API

Moving forward, we will want to pull in data from the CDWR telemetry API using an API key (located in the hidden creds folder). This will be done using the `get_telemetry_ts` function from the `cdssr` package. The following code snippet demonstrates how to pull in data for all telemetry stations within the Big Thompson watershed for a specified date range.


```{r}
startdate = "2026-03-01"
enddate = "2026-12-31"
#If you are going to pull in more than 2 years, you will want CDWR creds
creds <- read_yaml(here("creds/CDWRCreds.yml"))$api_key

all_data <- map2(telemetry$abbrev, telemetry$parameter, function(abbrev_oi, parameter_oi) {
    tryCatch({
      get_telemetry_ts(
        abbrev = abbrev_oi, 
        parameter = parameter_oi, 
        start_date = startdate, 
        end_date = enddate, 
        api_key = creds,
        timescale = "raw"
      )
    }, error = function(e) {
      message("Error pulling data for station: ", abbrev_oi, " and parameter: ", parameter_oi, " - ", e$message)
      return(NULL) # Explicitly return NULL on failure
    })
  }
) %>%
  bind_rows() %>%
  mutate(parameter = ifelse(parameter == "DISCHRG2", "DISCHRG", parameter))%>%
  mutate(meas_unit = ifelse(meas_unit == "CFS", "cfs", meas_unit))%>%
  left_join(
    telemetry %>%
    st_drop_geometry() %>%
    select(abbrev,station_name, station_type), by = "abbrev"
    )
```


```{r}
all_data%>%
  mutate(type_unit = paste0(station_type, " (", meas_unit, ")"))%>% 
  ggplot(.,
         aes(x = datetime, y = meas_value, color = type_unit)) +
  geom_line() +
  facet_wrap(~station_name, scales = "free_y") +
  labs(title = "Telemetry Data from Big Thompson Watershed",
       x = "Date",
       y = "Value",
       color = "Units")
```



## Trim out the sites we don't want

```{r}
telemetry%>%
  filter(station_status == "Active")%>%
  write_csv(here("data", "cdwr_sites_oi.csv"))

```

# Rewrite of `pull_wet_api` for loveland

Changes:
- Changed to use a csv file with the relevant site IDs and parameters to pull from the WET API
- Slight clean up of how data is manipulated using this CSV

```{r}

source("R/pull_wet_api_loveland.R")

```

## Testing new WET API function

```{r}
 sites <- c("reservoir_bypass", "sulzer", "cherry", "north_fork")

 new_WET_data <- map_dfr(sites,
                        ~pull_wet_api_loveland(
                          target_site = .x,
                          start_datetime = ymd_hms("2026-03-01 12:00:00", tz = "MST"),
                          end_datetime = Sys.time()
                        ))
 ggplot(new_WET_data, aes(x = DT_round_MT, y = value, color = site)) +
   geom_line() +
   facet_wrap(~ parameter, scales = "free_y") +
   labs(title = "WET Data from Big Thompson Watershed",
        x = "Date",
        y = "Value",
        color = "Parameter"   
 )
```

Seems like some funky stuff is going on with the WET API. Should be able to clean it up with `ross.wq.tools` a bit. BUT the API puller works and we can pull in the WET WQ data 

